Redefining CFRD Screening and Diagnosis in the Highly Effective Modulator Therapy Era (CF- SuGAR)

### 2025-09-16 Christine Chan/CF Registry/cfrd_1hr_ogtt_cgm_predictive.qmd
updated plot for Christine

```{r setup}
#| include: false
library(rspiro)
library(childsds)
library(tidyverse)
library(tableone)
library(pROC)
library(survival)
library(performance)
knitr::opts_knit$set(root.dir = "/Volumes/som/PEDS/RI Biostatistics Core/Shared/Shared Projects/Laura/Peds Endo")
```

```{r}
#| include: false
# read in separate data files
ogtt <- read.csv("./Christine Chan/1 hour OGTT in CFRD/Data_raw/ogtt.csv")
pft <- read.csv("./Christine Chan/1 hour OGTT in CFRD/Data_raw/pfts.csv")
a1c <- read.csv("./Christine Chan/1 hour OGTT in CFRD/Data_raw/a1c.csv")
demo <- read.csv("./Christine Chan/1 hour OGTT in CFRD/Data_raw/demographics.csv")
dob <- read.csv("./Christine Chan/1 hour OGTT in CFRD/Data_raw/dob.csv")
# merge dob and pfts
pft <- merge(pft, dob, by = "MRN", all.x = T, all.y = F)
pft <- unique(pft)
# checking for dups
# t <- ogtt[,c("MRN","FName","LName")]
# t <- unique(t)
# ct <- t %>% group_by(LName,FName) %>% count(unique(MRN))
# merge ogtt and demo
demo$FName <- NULL
demo$LName <- NULL
ogtt <- ogtt %>%
  group_by(MRN, Date.of.OGTT) %>%
  filter(row_number() == 1)
ogtt_demo <- merge(ogtt, demo, by = "MRN", all.x = T, all.y = T)
ogtt_demo$Date.of.OGTT <- mdy(ogtt_demo$Date.of.OGTT)
a1c$EncounterDate <- as.Date(a1c$Date.of.A1C, format = "%m/%d/%Y")
a1c$FName <- NULL
a1c$LName <- NULL
a1c$Date.of.OGTT <- NULL
a1c$Days.Between.A1C.and.OGTT <- NULL
ogtt_demo$EncounterDate <- ogtt_demo$Date.of.OGTT
ogtt_demo_a1c <- merge(ogtt_demo, a1c, by = c("MRN", "EncounterDate"), all.x = T, all.y = T)

@@

alldata <- merge(ogtt_demo_a1c, pft, by = c("MRN", "EncounterDate"), all.x = T, all.y = T)

@@

alldata$FEV1.pred <- pred_GLI(age = alldata$Age.at.PFT, height = alldata$ht_m, gender = alldata$sexnum, param = "FEV1")
alldata$FVC.pred <- pred_GLI(age = alldata$Age.at.PFT, height = alldata$ht_m, gender = alldata$sexnum, param = "FVC")
alldata$FEF.25.75.pred <- pred_GLI(age = alldata$Age.at.PFT, height = alldata$ht_m, gender = alldata$sexnum, param = "FEF2575")

@@

alldata$FEF.25.75.per.pred <- (alldata$FEF.25.75.Raw / alldata$FEF.25.75.pred) * 100

@@

alldata[alldata$MRN == 686476, ]$X1.hour.Glucose <- ifelse(alldata[alldata$MRN == 686476, ]$Age.At.Test..years. == 14.5, 185, alldata[alldata$MRN == 686476, ]$X1.hour.Glucose)

@@

alldata$class2 <- ifelse(alldata$Mutation.2 %in% c(
  "1717-1G->A", "1717-1G-A", "G542X", "Q493X", "R1162X", "R553X", "R553x", "W1089X", "W1282X", "1078delT", "R75X",
  "3659delC", "621+1G->T", "621+1G>T", "394delTT", "3120 + 1g->A", "1154InsTC", "1154insTC", "1213delT",
  "1259insA", "1288insTA", "3791delC", "E60X", "K710X", "2184delA", "CFTRdele2,3", "663delT", "Glu528",
  "1461ins4", "306insA", "R709X", "CFTRdele22-24", "2711delT", "2183AA- >G", "2183delAA>G"
), "I",
ifelse(alldata$Mutation.2 %in% c("F508", "F508del", "F508del ", "G85E", "I507", "I 507", "N1303K", "A559T", "R560T", "A561E"),

@@

  ifelse(alldata$Mutation.2 %in% c("G551D", "G551S", "S549N"), "III",
    ifelse(alldata$Mutation.2 %in% c("R334W", "R347P", "R117H", "R117C", "P67L", "L206W", "l206w", "I206w", "D614G", "R347H", "D1152H", "3849+10kbC>T"), "IV",
      ifelse(alldata$Mutation.2 %in% c(
        "A455E", "2789+5G->A", "3849+10CT", "3849+10kbC->T", "711+3A->G", "1898+5G->T",
        "P574H", "3272-26A->G5", "3272-26A->G"
      ), "V", "")

@@

alldata$severity2 <- ifelse(alldata$Mutation.2 %in% c(
  "1717-1G->A", "1717-1G-A", "G542X", "Q493X", "R1162X", "R553X", "R553x", "W1089X", "W1282X", "1078delT", "R75X",
  "3659delC", "621+1G->T", "621+1G>T", "394delTT", "3120 + 1g->A", "1154InsTC", "1154insTC", "1213delT",
  "1259insA", "1288insTA", "3791delC", "E60X", "K710X", "2184delA", "CFTRdele2,3", "663delT", "Glu528",
  "1461ins4", "306insA", "R709X", "CFTRdele22-24", "2711delT", "2183AA- >G", "2183delAA>G", "F508", "F508del", "F508del ", "G85E",
  "I507", "I 507", "N1303K", "A559T", "R560T", "A561E", "G551D", "G551S", "S549N", "1898+1G->A", "2143delT", "2183AA-G",
  "2183delAA->G", "2184insA", "S434X", "2585delT", "3120+1G->A", "3905insT", "406-1G->A", "712-1G->T", "L1245X", "Q2X", "R1158X",
  "R851X", "V520F", "L1077P", "S489X", "F311del", "W846X", "M1101K", "2622+1G->A", "S945L", "Y1092X", "R1066C", "G178R", "2957delT",
  "CFTRdele1", "W1204X", "q493x", "W57G", "1585-8G>A", "1717-8G->A", "G1244E"
), "Severe",
ifelse(alldata$Mutation.2 %in% c(
  "R334W", "R347P", "R117H", "R117C", "P67L", "L206W", "l206w", "I206w", "D614G", "R347H", "D1152H", "3849+10kbC>T", "D110H", "R1070W",
  "A455E", "2789+5G->A", "3849+10CT", "3849+10kbC->T", "711+3A->G", "1898+5G->T", "P574H", "3272-26A->G5", "3272-26A->G", "3849G->A", "H199Y", "H199y",
  "S492F", "S531P", "L1335P", "R352Q", "D1270N", "2789+2insA", "3849+10kbc->T", "Q98R", "Phe1052Val",
  "F1052V", "Tyr1014Cys", "Y1014C", "T338I", "5T", "c.2249C>T", "P750L"
), "Mild",
ifelse(alldata$Mutation.2 %in% c(
  "1262insA", "I506T", "3363delGT", "Q1209P", "3349insT", "G178R", "L467P", "317insC", "I336K",
  "ex14a", "3500-2A->G", "c.3407_3422del16", "3349insT", "1248+1G->A", "M470V", "1812-1G->A",
  "del X22-24", "Q237H", "Q237H", "W57G", "1078delA", "G178R", "2957delT", "G576A",
  "other", "Other", "Unknown", "unknown", "CFTRdele1", "1811+1.6kb A->", "UNK", "I336K", "M470V", "S1235R", "M470V", "F508C",
  "R1066H", "P205S"
), "Unk", "")

@@

alldata$class1 <- ifelse(alldata$Mutation.1 %in% c(
  "1717-1G->A", "1717-1G-A", "G542X", "Q493X", "R1162X", "R553X", "R553x", "W1089X", "W1282X", "1078delT", "R75X",
  "3659delC", "621+1G->T", "621+1G>T", "394delTT", "3120 + 1g->A", "1154InsTC", "1154insTC", "1213delT",
  "1259insA", "1288insTA", "3791delC", "E60X", "K710X", "2184delA", "CFTRdele2,3", "663delT", "Glu528",
  "1461ins4", "306insA", "R709X", "CFTRdele22-24", "2711delT", "2183AA- >G", "2183delAA>G"
), "I",
ifelse(alldata$Mutation.1 %in% c("F508", "F508del", "F508del ", "G85E", "I507", "I 507", "N1303K", "A559T", "R560T", "A561E"), "II",
  ifelse(alldata$Mutation.1 %in% c("G551D", "G551S", "S549N"), "III",
    ifelse(alldata$Mutation.1 %in% c("R334W", "R347P", "R117H", "R117C", "P67L", "L206W", "l206w", "I206w", "D614G", "R347H", "D1152H", "3849+10kbC>T"), "IV",
      ifelse(alldata$Mutation.1 %in% c(
        "A455E", "2789+5G->A", "3849+10CT", "3849+10kbC->T", "711+3A->G", "1898+5G->T",
        "P574H", "3272-26A->G5", "3272-26A->G"
      ), "V", "")

@@

alldata$severity1 <- ifelse(alldata$Mutation.1 %in% c(
  "1717-1G->A", "1717-1G-A", "G542X", "Q493X", "R1162X", "R553X", "R553x", "W1089X", "W1282X", "1078delT", "R75X",
  "3659delC", "621+1G->T", "621+1G>T", "394delTT", "3120 + 1g->A", "1154InsTC", "1154insTC", "1213delT",
  "1259insA", "1288insTA", "3791delC", "E60X", "K710X", "2184delA", "CFTRdele2,3", "663delT", "Glu528",
  "1461ins4", "306insA", "R709X", "CFTRdele22-24", "2711delT", "2183AA- >G", "2183delAA>G", "F508", "F508del", "F508del ", "G85E",
  "I507", "I 507", "N1303K", "A559T", "R560T", "A561E", "G551D", "G551S", "S549N", "1898+1G->A", "2143delT", "2183AA-G",
  "2183delAA->G", "2184insA", "S434X", "2585delT", "3120+1G->A", "3905insT", "406-1G->A", "712-1G->T", "L1245X", "Q2X", "R1158X",
  "R851X", "V520F", "L1077P", "S489X", "F311del", "W846X", "M1101K", "2622+1G->A", "S945L",
  "Y1092X", "R1066C", "G178R", "2957delT", "CFTRdele1", "W1204X", "q493x", "W57G", "1585-8G>A", "1717-8G->A", "G1244E"
), "Severe",
ifelse(alldata$Mutation.1 %in% c(
  "R334W", "R347P", "R117H", "R117C", "P67L", "L206W", "l206w", "I206w", "D614G", "R347H", "D1152H", "3849+10kbC>T", "D110H", "R1070W",
  "A455E", "2789+5G->A", "3849+10CT", "3849+10kbC->T", "711+3A->G", "1898+5G->T",
  "P574H", "3272-26A->G5", "3272-26A->G", "3849G->A", "H199Y", "H199y", "S492F", "S531P",
  "L1335P", "R352Q", "D1270N", "2789+2insA", "3849+10kbc->T", "Q98R", "Phe1052Val",
  "F1052V", "Tyr1014Cys", "Y1014C", "T338I", "5T", "c.2249C>T", "P750L"
), "Mild",
ifelse(alldata$Mutation.1 %in% c(
  "1262insA", "I506T", "3363delGT", "Q1209P", "3349insT", "G178R", "L467P", "317insC", "I336K",
  "ex14a", "3500-2A->G", "c.3407_3422del16", "3349insT", "1248+1G->A", "M470V", "1812-1G->A",
  "del X22-24", "Q237H", "Q237H", "W57G", "1078delA", "G178R", "2957delT", "G576A",
  "other", "Other", "Unknown", "unknown", "CFTRdele1", "1811+1.6kb A->", "UNK", "I336K", "M470V", "S1235R", "M470V", "F508C",
  "R1066H", "P205S"
), "Unk", "")

@@

alldata$severity2 <- ifelse(alldata$MRN %in% c(546678, 664400, 2015506), "Severe", alldata$severity2)
alldata$genoRisk <- ifelse(alldata$severity1 == "Mild" | alldata$severity2 == "Mild", "Low",
  ifelse(alldata$severity1 == "Severe" & alldata$severity2 == "Severe", "High", "")

@@

d <- alldata[, c("MRN", "FName", "LName", "Sex", "Race", "Latino.a.", "Mutation.1", "Mutation.2", "class1", "class2", "severity1", "severity2", "genoRisk")]

@@

col_check <- alldata[alldata$MRN %in% c(
  1006479, 1053930, 1078408, 1136992, 1289450, 1339023, 1413821, 1695512, 664400,
  686476, 707831, 712676, 739231, 739460, 787372, 800007, 848440, 860181, 875128, 897478, 905284, 908966,
  918124, 932695, 949042, 952250, 952252, 825118
), ]

@@

  "MRN", "FName", "LName", "Sex", "Race", "Latino.a.", "Mutation.1", "Mutation.2", "class1", "class2", "severity1", "severity2", "genoRisk",
  "Date of first OGTT", "Date", "Age.At.Test..years.", "Fasting.Glucose", "X1.hour.Glucose", "X2.hour.Glucose", "A1C",
  "DOB", "Age.at.PFT", "FVC.Raw", "Height..cm.", "Weight..kg.", "FEV1.Raw", "FEF.25.75.Raw", "ht_m",
  "sexnum", "FEV1.pred", "FVC.pred", "FEF.25.75.pred", "FEV1.per.pred", "FVC.per.pred", "FEF.25.75.per.pred", "BMI"

@@

alldata$race_eth <- ifelse(alldata$Latino.a. == "TRUE", "Hispanic", alldata$Race)

@@

alldata$age <- ifelse(is.na(alldata$Age.at.PFT), alldata$Age.At.Test..years., alldata$Age.at.PFT)

@@

alldata$bmip <- sds(alldata$BMI,

@@

  sex = alldata$Sex, male = "Male", female = "Female",

@@

alldata$cfrd_visit <- ifelse(is.na(alldata$X2.hour.Glucose), NA, ifelse(alldata$X2.hour.Glucose >= 200, 1, 0))

@@

cfrdbase <- alldata[alldata$Date == alldata$`Date of first OGTT` & alldata$cfrd_visit == 1, ]

@@

baseage <- alldata[alldata$Date == alldata$`Date of first OGTT`, c("MRN", "age")]

@@

basegluc <- alldata[alldata$Date == alldata$`Date of first OGTT`, c("MRN", "X1.hour.Glucose", "X2.hour.Glucose")]

@@

alldata$Baseline_1hr_glucose_155 <- ifelse(alldata$Baseline_1hr_glucose >= 155, 1, 0)
alldata$Baseline_1hr_glucose_140 <- ifelse(alldata$Baseline_1hr_glucose >= 140, 1, 0)

@@

alldata$Baseline_2hr_glucose_140 <- ifelse(alldata$Baseline_2hr_glucose >= 140, 1, 0)

@@

  !is.na(alldata$Fasting.Glucose) | !is.na(alldata$X1.hour.Glucose) | !is.na(alldata$X2.hour.Glucose),

@@

mod <- read.csv("./Christine Chan/1 hour OGTT in CFRD/Data_raw/modulators.csv", na.strings = c("", " ", NA))

@@

    as.Date(pmin(mod[i, ]$Kalydeco.Start.Date, mod[i, ]$Orkambi.Start.Date, mod[i, ]$Symdeko.Start.Date, mod[i, ]$Trikafta.Start.Date, na.rm = T))
  mod[i, ]$he_modstart <- as.Date(pmin(mod[i, ]$Kalydeco.Start.Date, mod[i, ]$Trikafta.Start.Date, na.rm = T))
  mod[i, ]$me_modstart <- as.Date(pmin(mod[i, ]$Orkambi.Start.Date, mod[i, ]$Symdeko.Start.Date, na.rm = T))

@@

alldata$onmod <- ifelse(!is.na(alldata$modstart) & alldata$modstart <= alldata$Date, 1,

@@

alldata$onmod <- ifelse(is.na(alldata$modstart) & is.na(alldata$onmod), 0, alldata$onmod)
alldata$onhemod <- ifelse(!is.na(alldata$he_modstart) & alldata$he_modstart <= alldata$Date, 1,
  ifelse(!is.na(alldata$he_modstart) & alldata$he_modstart > alldata$Date, 0, NA)

@@

alldata$onhemod <- ifelse(is.na(alldata$he_modstart) & is.na(alldata$onhemod), 0, alldata$onhemod)
alldata$onmemod <- ifelse(!is.na(alldata$me_modstart) & alldata$me_modstart <= alldata$Date, 1,
  ifelse(!is.na(alldata$me_modstart) & alldata$me_modstart > alldata$Date, 0, NA)

@@

alldata$onmemod <- ifelse(is.na(alldata$me_modstart) & is.na(alldata$onmemod), 0, alldata$onmemod)
alldata$FEV.1.per.pred_FVC.per.pred <- alldata$FEV1.per.pred / alldata$FVC.per.pred

>>>>

```{r setup}
#| include: false
library(rspiro)
library(childsds)
library(tidyverse)
library(tableone)
library(pROC)
library(survival)
library(performance)
knitr::opts_knit$set(
  root.dir = "/Volumes/som/PEDS/RI Biostatistics Core/Shared/Shared Projects/Laura/Peds Endo"
)
```

```{r}
#| include: false
# read in separate data files
ogtt <- read.csv("./Christine Chan/1 hour OGTT in CFRD/Data_raw/ogtt.csv")
pft <- read.csv("./Christine Chan/1 hour OGTT in CFRD/Data_raw/pfts.csv")
a1c <- read.csv("./Christine Chan/1 hour OGTT in CFRD/Data_raw/a1c.csv")
demo <- read.csv(
  "./Christine Chan/1 hour OGTT in CFRD/Data_raw/demographics.csv"
)
dob <- read.csv("./Christine Chan/1 hour OGTT in CFRD/Data_raw/dob.csv")
# merge dob and pfts
pft <- merge(pft, dob, by = "MRN", all.x = T, all.y = F)
pft <- unique(pft)
# checking for dups
# t <- ogtt[,c("MRN","FName","LName")]
# t <- unique(t)
# ct <- t %>% group_by(LName,FName) %>% count(unique(MRN))
# merge ogtt and demo
demo$FName <- NULL
demo$LName <- NULL
ogtt <- ogtt %>%
  group_by(MRN, Date.of.OGTT) %>%
  filter(row_number() == 1)
ogtt_demo <- merge(ogtt, demo, by = "MRN", all.x = T, all.y = T)
ogtt_demo$Date.of.OGTT <- mdy(ogtt_demo$Date.of.OGTT)
a1c$EncounterDate <- as.Date(a1c$Date.of.A1C, format = "%m/%d/%Y")
a1c$FName <- NULL
a1c$LName <- NULL
a1c$Date.of.OGTT <- NULL
a1c$Days.Between.A1C.and.OGTT <- NULL
ogtt_demo$EncounterDate <- ogtt_demo$Date.of.OGTT
ogtt_demo_a1c <- merge(
  ogtt_demo,
  a1c,
  by = c("MRN", "EncounterDate"),
  all.x = T,
  all.y = T
)

@@

alldata <- merge(
  ogtt_demo_a1c,
  pft,
  by = c("MRN", "EncounterDate"),
  all.x = T,
  all.y = T

@@

alldata$FEV1.pred <- pred_GLI(
  age = alldata$Age.at.PFT,
  height = alldata$ht_m,
  gender = alldata$sexnum,
  param = "FEV1"
alldata$FVC.pred <- pred_GLI(
  age = alldata$Age.at.PFT,
  height = alldata$ht_m,
  gender = alldata$sexnum,
  param = "FVC"
alldata$FEF.25.75.pred <- pred_GLI(
  age = alldata$Age.at.PFT,
  height = alldata$ht_m,
  gender = alldata$sexnum,
  param = "FEF2575"

@@

alldata$FEF.25.75.per.pred <- (alldata$FEF.25.75.Raw / alldata$FEF.25.75.pred) *
  100

@@

alldata[alldata$MRN == 686476, ]$X1.hour.Glucose <- ifelse(
  alldata[alldata$MRN == 686476, ]$Age.At.Test..years. == 14.5,
  185,
  alldata[alldata$MRN == 686476, ]$X1.hour.Glucose

@@

alldata$class2 <- ifelse(
  alldata$Mutation.2 %in%
    c(
      "1717-1G->A",
      "1717-1G-A",
      "G542X",
      "Q493X",
      "R1162X",
      "R553X",
      "R553x",
      "W1089X",
      "W1282X",
      "1078delT",
      "R75X",
      "3659delC",
      "621+1G->T",
      "621+1G>T",
      "394delTT",
      "3120 + 1g->A",
      "1154InsTC",
      "1154insTC",
      "1213delT",
      "1259insA",
      "1288insTA",
      "3791delC",
      "E60X",
      "K710X",
      "2184delA",
      "CFTRdele2,3",
      "663delT",
      "Glu528",
      "1461ins4",
      "306insA",
      "R709X",
      "CFTRdele22-24",
      "2711delT",
      "2183AA- >G",
      "2183delAA>G"
    ),
  "I",
  ifelse(
    alldata$Mutation.2 %in%
      c(
        "F508",
        "F508del",
        "F508del ",
        "G85E",
        "I507",
        "I 507",
        "N1303K",
        "A559T",
        "R560T",
        "A561E"
      ),

@@

    ifelse(
      alldata$Mutation.2 %in% c("G551D", "G551S", "S549N"),
      "III",
      ifelse(
        alldata$Mutation.2 %in%
          c(
            "R334W",
            "R347P",
            "R117H",
            "R117C",
            "P67L",
            "L206W",
            "l206w",
            "I206w",
            "D614G",
            "R347H",
            "D1152H",
            "3849+10kbC>T"
          ),
        "IV",
        ifelse(
          alldata$Mutation.2 %in%
            c(
              "A455E",
              "2789+5G->A",
              "3849+10CT",
              "3849+10kbC->T",
              "711+3A->G",
              "1898+5G->T",
              "P574H",
              "3272-26A->G5",
              "3272-26A->G"
            ),
          "V",
          ""

@@

alldata$severity2 <- ifelse(
  alldata$Mutation.2 %in%
    c(
      "1717-1G->A",
      "1717-1G-A",
      "G542X",
      "Q493X",
      "R1162X",
      "R553X",
      "R553x",
      "W1089X",
      "W1282X",
      "1078delT",
      "R75X",
      "3659delC",
      "621+1G->T",
      "621+1G>T",
      "394delTT",
      "3120 + 1g->A",
      "1154InsTC",
      "1154insTC",
      "1213delT",
      "1259insA",
      "1288insTA",
      "3791delC",
      "E60X",
      "K710X",
      "2184delA",
      "CFTRdele2,3",
      "663delT",
      "Glu528",
      "1461ins4",
      "306insA",
      "R709X",
      "CFTRdele22-24",
      "2711delT",
      "2183AA- >G",
      "2183delAA>G",
      "F508",
      "F508del",
      "F508del ",
      "G85E",
      "I507",
      "I 507",
      "N1303K",
      "A559T",
      "R560T",
      "A561E",
      "G551D",
      "G551S",
      "S549N",
      "1898+1G->A",
      "2143delT",
      "2183AA-G",
      "2183delAA->G",
      "2184insA",
      "S434X",
      "2585delT",
      "3120+1G->A",
      "3905insT",
      "406-1G->A",
      "712-1G->T",
      "L1245X",
      "Q2X",
      "R1158X",
      "R851X",
      "V520F",
      "L1077P",
      "S489X",
      "F311del",
      "W846X",
      "M1101K",
      "2622+1G->A",
      "S945L",
      "Y1092X",
      "R1066C",
      "G178R",
      "2957delT",
      "CFTRdele1",
      "W1204X",
      "q493x",
      "W57G",
      "1585-8G>A",
      "1717-8G->A",
      "G1244E"
    ),
  "Severe",
  ifelse(
    alldata$Mutation.2 %in%
      c(
        "R334W",
        "R347P",
        "R117H",
        "R117C",
        "P67L",
        "L206W",
        "l206w",
        "I206w",
        "D614G",
        "R347H",
        "D1152H",
        "3849+10kbC>T",
        "D110H",
        "R1070W",
        "A455E",
        "2789+5G->A",
        "3849+10CT",
        "3849+10kbC->T",
        "711+3A->G",
        "1898+5G->T",
        "P574H",
        "3272-26A->G5",
        "3272-26A->G",
        "3849G->A",
        "H199Y",
        "H199y",
        "S492F",
        "S531P",
        "L1335P",
        "R352Q",
        "D1270N",
        "2789+2insA",
        "3849+10kbc->T",
        "Q98R",
        "Phe1052Val",
        "F1052V",
        "Tyr1014Cys",
        "Y1014C",
        "T338I",
        "5T",
        "c.2249C>T",
        "P750L"
      ),
    "Mild",
    ifelse(
      alldata$Mutation.2 %in%
        c(
          "1262insA",
          "I506T",
          "3363delGT",
          "Q1209P",
          "3349insT",
          "G178R",
          "L467P",
          "317insC",
          "I336K",
          "ex14a",
          "3500-2A->G",
          "c.3407_3422del16",
          "3349insT",
          "1248+1G->A",
          "M470V",
          "1812-1G->A",
          "del X22-24",
          "Q237H",
          "Q237H",
          "W57G",
          "1078delA",
          "G178R",
          "2957delT",
          "G576A",
          "other",
          "Other",
          "Unknown",
          "unknown",
          "CFTRdele1",
          "1811+1.6kb A->",
          "UNK",
          "I336K",
          "M470V",
          "S1235R",
          "M470V",
          "F508C",
          "R1066H",
          "P205S"
        ),
      "Unk",
      ""

@@

alldata$class1 <- ifelse(
  alldata$Mutation.1 %in%
    c(
      "1717-1G->A",
      "1717-1G-A",
      "G542X",
      "Q493X",
      "R1162X",
      "R553X",
      "R553x",
      "W1089X",
      "W1282X",
      "1078delT",
      "R75X",
      "3659delC",
      "621+1G->T",
      "621+1G>T",
      "394delTT",
      "3120 + 1g->A",
      "1154InsTC",
      "1154insTC",
      "1213delT",
      "1259insA",
      "1288insTA",
      "3791delC",
      "E60X",
      "K710X",
      "2184delA",
      "CFTRdele2,3",
      "663delT",
      "Glu528",
      "1461ins4",
      "306insA",
      "R709X",
      "CFTRdele22-24",
      "2711delT",
      "2183AA- >G",
      "2183delAA>G"
    ),
  "I",
  ifelse(
    alldata$Mutation.1 %in%
      c(
        "F508",
        "F508del",
        "F508del ",
        "G85E",
        "I507",
        "I 507",
        "N1303K",
        "A559T",
        "R560T",
        "A561E"
      ),
    ifelse(
      alldata$Mutation.1 %in% c("G551D", "G551S", "S549N"),
      "III",
      ifelse(
        alldata$Mutation.1 %in%
          c(
            "R334W",
            "R347P",
            "R117H",
            "R117C",
            "P67L",
            "L206W",
            "l206w",
            "I206w",
            "D614G",
            "R347H",
            "D1152H",
            "3849+10kbC>T"
          ),
        "IV",
        ifelse(
          alldata$Mutation.1 %in%
            c(
              "A455E",
              "2789+5G->A",
              "3849+10CT",
              "3849+10kbC->T",
              "711+3A->G",
              "1898+5G->T",
              "P574H",
              "3272-26A->G5",
              "3272-26A->G"
            ),
          "V",
          ""

@@

alldata$severity1 <- ifelse(
  alldata$Mutation.1 %in%
    c(
      "1717-1G->A",
      "1717-1G-A",
      "G542X",
      "Q493X",
      "R1162X",
      "R553X",
      "R553x",
      "W1089X",
      "W1282X",
      "1078delT",
      "R75X",
      "3659delC",
      "621+1G->T",
      "621+1G>T",
      "394delTT",
      "3120 + 1g->A",
      "1154InsTC",
      "1154insTC",
      "1213delT",
      "1259insA",
      "1288insTA",
      "3791delC",
      "E60X",
      "K710X",
      "2184delA",
      "CFTRdele2,3",
      "663delT",
      "Glu528",
      "1461ins4",
      "306insA",
      "R709X",
      "CFTRdele22-24",
      "2711delT",
      "2183AA- >G",
      "2183delAA>G",
      "F508",
      "F508del",
      "F508del ",
      "G85E",
      "I507",
      "I 507",
      "N1303K",
      "A559T",
      "R560T",
      "A561E",
      "G551D",
      "G551S",
      "S549N",
      "1898+1G->A",
      "2143delT",
      "2183AA-G",
      "2183delAA->G",
      "2184insA",
      "S434X",
      "2585delT",
      "3120+1G->A",
      "3905insT",
      "406-1G->A",
      "712-1G->T",
      "L1245X",
      "Q2X",
      "R1158X",
      "R851X",
      "V520F",
      "L1077P",
      "S489X",
      "F311del",
      "W846X",
      "M1101K",
      "2622+1G->A",
      "S945L",
      "Y1092X",
      "R1066C",
      "G178R",
      "2957delT",
      "CFTRdele1",
      "W1204X",
      "q493x",
      "W57G",
      "1585-8G>A",
      "1717-8G->A",
      "G1244E"
    ),
  "Severe",
  ifelse(
    alldata$Mutation.1 %in%
      c(
        "R334W",
        "R347P",
        "R117H",
        "R117C",
        "P67L",
        "L206W",
        "l206w",
        "I206w",
        "D614G",
        "R347H",
        "D1152H",
        "3849+10kbC>T",
        "D110H",
        "R1070W",
        "A455E",
        "2789+5G->A",
        "3849+10CT",
        "3849+10kbC->T",
        "711+3A->G",
        "1898+5G->T",
        "P574H",
        "3272-26A->G5",
        "3272-26A->G",
        "3849G->A",
        "H199Y",
        "H199y",
        "S492F",
        "S531P",
        "L1335P",
        "R352Q",
        "D1270N",
        "2789+2insA",
        "3849+10kbc->T",
        "Q98R",
        "Phe1052Val",
        "F1052V",
        "Tyr1014Cys",
        "Y1014C",
        "T338I",
        "5T",
        "c.2249C>T",
        "P750L"
      ),
    "Mild",
    ifelse(
      alldata$Mutation.1 %in%
        c(
          "1262insA",
          "I506T",
          "3363delGT",
          "Q1209P",
          "3349insT",
          "G178R",
          "L467P",
          "317insC",
          "I336K",
          "ex14a",
          "3500-2A->G",
          "c.3407_3422del16",
          "3349insT",
          "1248+1G->A",
          "M470V",
          "1812-1G->A",
          "del X22-24",
          "Q237H",
          "Q237H",
          "W57G",
          "1078delA",
          "G178R",
          "2957delT",
          "G576A",
          "other",
          "Other",
          "Unknown",
          "unknown",
          "CFTRdele1",
          "1811+1.6kb A->",
          "UNK",
          "I336K",
          "M470V",
          "S1235R",
          "M470V",
          "F508C",
          "R1066H",
          "P205S"
        ),
      "Unk",
      ""

@@

alldata$severity2 <- ifelse(
  alldata$MRN %in% c(546678, 664400, 2015506),
  "Severe",
  alldata$severity2
alldata$genoRisk <- ifelse(
  alldata$severity1 == "Mild" | alldata$severity2 == "Mild",
  "Low",
  ifelse(
    alldata$severity1 == "Severe" & alldata$severity2 == "Severe",
    "High",
    ""

@@

d <- alldata[, c(
  "MRN",
  "FName",
  "LName",
  "Sex",
  "Race",
  "Latino.a.",
  "Mutation.1",
  "Mutation.2",
  "class1",
  "class2",
  "severity1",
  "severity2",
  "genoRisk"
)]

@@

col_check <- alldata[
  alldata$MRN %in%
    c(
      1006479,
      1053930,
      1078408,
      1136992,
      1289450,
      1339023,
      1413821,
      1695512,
      664400,
      686476,
      707831,
      712676,
      739231,
      739460,
      787372,
      800007,
      848440,
      860181,
      875128,
      897478,
      905284,
      908966,
      918124,
      932695,
      949042,
      952250,
      952252,
      825118
    ),

@@

  "MRN",
  "FName",
  "LName",
  "Sex",
  "Race",
  "Latino.a.",
  "Mutation.1",
  "Mutation.2",
  "class1",
  "class2",
  "severity1",
  "severity2",
  "genoRisk",
  "Date of first OGTT",
  "Date",
  "Age.At.Test..years.",
  "Fasting.Glucose",
  "X1.hour.Glucose",
  "X2.hour.Glucose",
  "A1C",
  "DOB",
  "Age.at.PFT",
  "FVC.Raw",
  "Height..cm.",
  "Weight..kg.",
  "FEV1.Raw",
  "FEF.25.75.Raw",
  "ht_m",
  "sexnum",
  "FEV1.pred",
  "FVC.pred",
  "FEF.25.75.pred",
  "FEV1.per.pred",
  "FVC.per.pred",
  "FEF.25.75.per.pred",
  "BMI"

@@

alldata$race_eth <- ifelse(
  alldata$Latino.a. == "TRUE",
  "Hispanic",
  alldata$Race

@@

alldata$age <- ifelse(
  is.na(alldata$Age.at.PFT),
  alldata$Age.At.Test..years.,
  alldata$Age.at.PFT

@@

alldata$bmip <- sds(
  alldata$BMI,

@@

  sex = alldata$Sex,
  male = "Male",
  female = "Female",

@@

alldata$cfrd_visit <- ifelse(
  is.na(alldata$X2.hour.Glucose),
  NA,
  ifelse(alldata$X2.hour.Glucose >= 200, 1, 0)

@@

cfrdbase <- alldata[
  alldata$Date == alldata$`Date of first OGTT` & alldata$cfrd_visit == 1,

@@

baseage <- alldata[
  alldata$Date == alldata$`Date of first OGTT`,
  c("MRN", "age")

@@

basegluc <- alldata[
  alldata$Date == alldata$`Date of first OGTT`,
  c("MRN", "X1.hour.Glucose", "X2.hour.Glucose")

@@

alldata$Baseline_1hr_glucose_155 <- ifelse(
  alldata$Baseline_1hr_glucose >= 155,
  1,
  0
alldata$Baseline_1hr_glucose_140 <- ifelse(
  alldata$Baseline_1hr_glucose >= 140,
  1,
  0

@@

alldata$Baseline_2hr_glucose_140 <- ifelse(
  alldata$Baseline_2hr_glucose >= 140,
  1,
  0

@@

  !is.na(alldata$Fasting.Glucose) |
    !is.na(alldata$X1.hour.Glucose) |
    !is.na(alldata$X2.hour.Glucose),

@@

mod <- read.csv(
  "./Christine Chan/1 hour OGTT in CFRD/Data_raw/modulators.csv",
  na.strings = c("", " ", NA)

@@

    as.Date(pmin(
      mod[i, ]$Kalydeco.Start.Date,
      mod[i, ]$Orkambi.Start.Date,
      mod[i, ]$Symdeko.Start.Date,
      mod[i, ]$Trikafta.Start.Date,
      na.rm = T
    ))
  mod[i, ]$he_modstart <- as.Date(pmin(
    mod[i, ]$Kalydeco.Start.Date,
    mod[i, ]$Trikafta.Start.Date,
    na.rm = T
  ))
  mod[i, ]$me_modstart <- as.Date(pmin(
    mod[i, ]$Orkambi.Start.Date,
    mod[i, ]$Symdeko.Start.Date,
    na.rm = T
  ))

@@

alldata$onmod <- ifelse(
  !is.na(alldata$modstart) & alldata$modstart <= alldata$Date,
  1,

@@

alldata$onmod <- ifelse(
  is.na(alldata$modstart) & is.na(alldata$onmod),
  0,
  alldata$onmod

@@

alldata$onhemod <- ifelse(
  !is.na(alldata$he_modstart) & alldata$he_modstart <= alldata$Date,
  1,
  ifelse(
    !is.na(alldata$he_modstart) & alldata$he_modstart > alldata$Date,
    0,
    NA

@@

alldata$onhemod <- ifelse(
  is.na(alldata$he_modstart) & is.na(alldata$onhemod),
  0,
  alldata$onhemod
alldata$onmemod <- ifelse(
  !is.na(alldata$me_modstart) & alldata$me_modstart <= alldata$Date,
  1,
  ifelse(
    !is.na(alldata$me_modstart) & alldata$me_modstart > alldata$Date,
    0,
    NA
alldata$onmemod <- ifelse(
